\Huge Conferences in the Era of Expensive Carbon

### 2019-06-21 main.tex
Checkpoint

\title{\Huge Conferences in an Era of Expensive Carbon}
% \title{\Huge ACM Conferences and Cost of Carbon}
\author{Benjamin C. Pierce (University of Pennsylvania) \\
Michael Hicks (University of Maryland) \\
Crista Lopes (University California, Irvine) \\
Jens Palsberg (University of California, Los Angeles)}

@@

How should ACM respond to these facts? We are the members of SIGPLAN's
(the special interest group on programming languages) Climate
Committee,\footnote{https://www.sigplan.org/Resources/Climate/} and we
have been researching this 
question for the past couple of years. Our work has led to two concrete
proposals. First, all ACM conferences should publicly {\em account} for CO2e
emitted as a result of putting them on. Second, ACM should put a
\emph{price} on carbon in conference budgets, to create pressure on
organizers to reduce their footprints.

\SECTION{Mandate public accountability.}  
% \bcp{The section headers could use some wordsmithing!}

Our first proposal is simple: Every ACM-sponsored conference should
calculate and publicly report its carbon footprint (both aggregate and
per-participant). Most conferences' footprints will be dominated by
participants air travel, but the accounting should go beyond this to include
air travel to in-person PC meetings and estimated hotel and food CO2e
footprints.

@@

SIGPLAN has been considering how to reduce CO2e usage over the past couple
of years, informed by an accounting of its CO2e expenditures. This
discussion has led the organizers of some of its flagship
conferences\mwh{name which conferences}\bcp{POPL and ICFP, not PLDI; I don't
  remember where SPLASH stands} to switch from in-person to on-line program
committee meetings, joining a trend among other SIGs. \mwh{Mention Jens
  survey here.} It has also prompted conferences to increase their
investments in livestreaming and talk recording to permit remote
participation.

SIGPLAN has also been exploring how to mitigate a conference's CO2e emissions
by purchasing \emph{carbon
  offsets}~\cite{CarbonOFfsetReport}.
%
A carbon offset is sold by a vendor who uses the raised funds to finance an
activity that permanently avoids emitting some amount of greenhouse gases.
The veracity and permanence of this CO2e emission avoidance is certified by
a watchdog organization. (Typically, planting trees is not a certifiable
activity, since it's hard to guarantee the trees won't be cut down, but
installing methane capture devices on landfill sites or buying
fuel-efficient stoves to replace open-fire cooking in poor communities can
be.)
%
Many organizations, including companies such as Google, Dell, Microsoft,
General Motors, Delta Airlines, Lyft, and Expedia, numerous universities,
and some academic societies, now buy carbon offsets in order to reduce their
net footprint. ACM conferences can do the same, and the purchases can be
included in the public accounting we are proposing.  In fact, ACM recently
decided that conferences could offer attendees the option to pay for an
offset during registration; the conference then uses the designated funds to
purchase the offset from a qualified vendor.  Some conferences are also
experimenting with corporate sponsorship of bulk offsets for the whole
conference.\footnote{We acknowledge that carbon offsets are a
  controversial tool\bcp{Cite the thing the reviewer pointed us to, as well
    as the recent controversial UN report.}.  Besides issues with
  verifiability and permanence, some observers worry that---especially given
  their currently artificially-low cost---they will encourage complacency,
  leading organizations to postpone making the difficult changes that will
  be required to sustainably decarbonize society.  We believe that offsets
  are a short-term measure, to be considered in parallel with, not instead
  of, more difficult cutting.}

Beyond encouraging conferences to take these sorts of simple steps, public
accounting of conference emissions footprints will generate valuable data
that future organizers can use to optimize their own footprints.
%
\bcp{Talk about the little bit of data gathering that we've done.  Make the
  point that, if the public accounting of conference carbon footprints is
  detailed enough (e.g., including an origin city for each participant,
  rather than just aggregate statistics on average per-participant
  footprint), and if it's publicized in a centralized place and in a uniform
  format, then it becomes a resource for all sorts of useful analysis.}

ACM can provide tools for gathering and publicizing this data. As a start,
SIGPLAN recently
worked with ACM's SIG Governing Board to develop an air-travel-focused CO2e
calculator for conferences. Users can upload conference registration data,
and the calculator will estimate the CO2e cost of air travel for those
sufficiently far away. This calculator could be expanded to account for
other conference emissions.

@@

\begin{figure}[t]
\centering
\includegraphics[width=6in]{SIGPLAN-confs.png}
\includegraphics[width=6in]{ParticipantsOriginAll.png}
\caption{Top: carbon footprint per participant for travel to recent SIGPLAN
  conferences\iflater\bcp{Get better file from the other paper, or get
    better version of this one from Crista.}\fi.  Bottom: Origin of
  participants for each conference.  \bcp{Explain in more detail, once we
    have Crista's new figures.}} 
\label{fig:confs}
\end{figure}

The registration data we've gathered from SIGPLAN conferences illustrates
some of the difficulties.  
%
\bcp{Explain the two diagrams.  Discuss how they can be used to draw
  contradictory conclusions.  E.g., one of the many footprint-reduction
  measures that have been discussed in SIGPLAN~\cite{ClimateCommitteReport}
  is locating conferences near their historical ``center of mass.''  Looking
  at the top diagram, one might thus be tempted to locate all conferences in
  the northeast US or western Europe.  But the bottom figure shows that much
  of the participation in each instance of a conference is local; always
  holding a conference at its ``center of mass'' will damage local
  communities everywhere else.  (Maybe this would be a good place to refer
  to~\cite{ClimateCommitteReport}---we can enumerate some of the
  alternatives that one might have hoped the data would help us choose
  between.)}  
%
\bcp{Maybe we can replace the bottom part of the figure by some clearer way
  of visualizing the observation that much of the attendance at each
  conference is local?  E.g., what if, instead of color-coding the different
  sections of the bars by continent, we instead color-coded them by distance
  traveled (in buckets of 500 miles, say)?}

>>>>

\title{\Huge Conferences in the Era of Expensive Carbon}
% \title{\Huge ACM Conferences and Cost of Carbon}
\author{Benjamin C. Pierce (University of Pennsylvania) \\
Michael Hicks (University of Maryland) \\
Crista Lopes (University California, Irvine) \\
Jens Palsberg (University of California, Los Angeles)}

@@

How should ACM respond to these facts?  Two years ago, SIGPLAN (ACM's
Special Interest Group on Programming Languages) convened an ad hoc Climate
Committee\footnote{https://www.sigplan.org/Resources/Climate/} to research
this question.  After investigating many
options~\cite{ClimateCommitteReport} and discussing intensively with SIGPLAN
members and conference organizers, ACM staff, and other SIG leaders, we have
two concrete proposals. First, all ACM conferences should publicly {\em
  account} for CO2e emitted as a result of putting them on. Second, ACM
should put a \emph{price} on carbon in conference budgets, to create
pressure on organizers to reduce their footprints.

\SECTION{Mandate public accountability.}

Our first proposal is simple: Every ACM-sponsored conference should
calculate and publicly report its carbon footprint. Most conferences'
footprints will be dominated by participants air travel, but the accounting
should go beyond this to include air travel to in-person PC meetings and
estimated hotel and food CO2e footprints.

@@

SIGPLAN has been considering how to reduce CO2e usage over the past couple
of years, informed by an accounting of its CO2e expenditures. This
discussion has led the organizers of two of its flagship conferences (POPL
and ICFP) to switch from in-person to on-line program committee meetings,
joining a trend among other SIGs (\bcp{ZZZ} out of \bcp{WWW} report that
their flagship meetings now have virtual PC meetings). It has also prompted
conferences to increase their investments in livestreaming and talk
recording to permit remote participation.

SIGPLAN has also explored how to mitigate a conference's CO2e emissions by
purchasing \emph{carbon offsets}~\cite{CarbonOFfsetReport}.
%
A carbon offset is sold by a vendor who uses the raised funds to finance an
activity that permanently avoids emitting some amount of greenhouse gases.
The veracity and permanence of this CO2e emission avoidance is certified by
a watchdog organization. (Typically, planting trees is not a certifiable
activity, since it's hard to guarantee the trees won't be cut down, but
installing methane capture devices on landfill sites or buying
fuel-efficient stoves to replace open-fire cooking in poor communities can
be.)
%
Many organizations, including companies such as Google, Dell, Microsoft,
General Motors, Delta Airlines, Lyft, and Expedia, numerous universities,
and some academic societies, now buy carbon offsets in order to reduce their
net footprint. ACM conferences can do the same, and the purchases can be
included in the public accounting we are proposing.  In fact, ACM recently
decided that conferences could offer attendees the option to pay for an
offset during registration; the conference then uses the designated funds to
purchase the offset from a qualified vendor.  Some conferences are also
experimenting with corporate sponsorship of bulk offsets for the whole
conference.\footnote{We acknowledge that carbon offsets are a controversial
  tool~\cite{anderson2012inconvenient,
  carbon-offsets-are-not-our-get-out-jail-free-card}.  Besides issues with
  verifiability and permanence, some observers worry that---especially given
  their currently artificially-low cost---they will encourage complacency,
  leading organizations to postpone making the difficult changes that will
  be required to sustainably decarbonize society.  We believe that offsets
  are a short-term measure, to be considered in parallel with, not instead
  of, more difficult cutting.}

Beyond encouraging conferences to take these sorts of simple steps, public
accounting of conference emissions footprints will generate valuable data
that future organizers can use to optimize their own footprints.  Emissions
data should be published both in aggregate form (total footprint for the
conference and per-participant footprint), for quick reference, and in a raw
form (a suitably anonymized record for each individual attending each
conference, including originating city), and it should be made available in
a central place and in a uniform format, to facilitate analysis.

ACM can provide tools for gathering and publicizing this data. As a start,
SIGPLAN recently worked with ACM's SIG Governing Board to develop an
air-travel-focused CO2e calculator for conferences. Users can upload
conference registration data, and the calculator will estimate the CO2e cost
of air travel for those sufficiently far away. This calculator could be
expanded to account for other forms of conference emissions.

@@

\begin{figure}[t]
\centering
\includegraphics[width=6in]{SIGPLAN-confs2.png}
\includegraphics[width=6in]{ParticipantsOriginAll2.png}
% \includegraphics[width=6in]{Distances.png}
\caption{Top: Carbon footprint per participant for travel to recent SIGPLAN
  conferences\iflater\bcp{Get better file from the other paper, or get
    better version of this one from Crista.}\fi.  (Small dots represent
  \bcp{XXX} tons of CO2e per participant; the largest dot represents
  \bcp{YYY} tons per participant.) Bottom: Breakdown of 
  continent-of-origin for participants in each conference. 
% Bottom: Distances traveled by participants, one way, in Km; middle dashed
% line is the median travel distance. 
}
\label{fig:confs}
\end{figure}

Registration data from recent SIGPLAN conferences illustrates
some of the difficulties.  
%
\bcp{Explain the two diagrams.  Discuss how they can be used to draw
  contradictory conclusions.  E.g., one of the many footprint-reduction
  measures that have been discussed in SIGPLAN~\cite{ClimateCommitteReport}
  is locating conferences near their historical ``center of mass.''  Looking
  at the top diagram, one might thus be tempted to locate all conferences in
  the northeast US or western Europe.  But the bottom figure shows that much
  of the participation in each instance of a conference is local; always
  holding a conference at its ``center of mass'' will damage local
  communities everywhere else.  (Maybe this would be a good place to refer
  to~\cite{ClimateCommitteReport}---we can enumerate some of the
  alternatives that one might have hoped the data would help us choose
  between.)}  
%
\bcp{Maybe we can replace the bottom part of the figure by some clearer way
  of visualizing the observation that much of the attendance at each
  conference is local?  E.g., what if, instead of color-coding the different
  sections of the bars by continent, we instead color-coded them by distance
  traveled (in buckets of 500 miles, say)?}
